---
title: "European Values Study 2017 Report"
subtitle: "Dynamic Report for SURV675 Final Project"
author: "Annia Mejia"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    number_sections: true
params:
  country:  "overall"
  outcome:  "v72"
  controls: !r character(0)
  age_poly: 1
  data:     NULL
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 8,
  fig.height = 4.5,
  dpi = 120
)

library(tidyverse)
library(broom)
library(kableExtra)
```

```{r data-prep}
load_evs_data <- function(path = "data/evs_data.rds") {
  if (!file.exists(path)) stop("EVS data file not found. Check the path.")
  readRDS(path)
}

filter_data <- function(data, country_input) {
  if (country_input == "overall") return(data)
  data %>% filter(country == country_input)
}

build_formula <- function(outcome, controls, age_poly) {
  age_term <- if (age_poly == 1) "age" else paste0("poly(age, ", age_poly, ")")
  rhs      <- paste(c(age_term, controls), collapse = " + ")
  paste(outcome, "~", rhs)
}

fit_model <- function(data, outcome, controls, age_poly) {
  f <- as.formula(build_formula(outcome, controls, age_poly))
  lm(f, data = data, na.action = na.omit)
}

outcome_label <- function(v) {
  switch(v,
    v72 = "Gender attitude (v72)",
    v80 = "Immigration attitude (v80)",
    v
  )
}

# Unpack parameters passed from Shiny
country_param  <- params$country
outcome_param  <- params$outcome
controls_param <- if (is.null(params$controls)) character(0) else params$controls
age_poly_param <- params$age_poly

country_disp  <- if (country_param == "overall") "All countries (overall sample)" else country_param
formula_str   <- build_formula(outcome_param, controls_param, age_poly_param)
controls_disp <- if (length(controls_param) == 0) "None" else paste(controls_param, collapse = ", ")

df <- if (!is.null(params$data)) {
  params$data
} else {
  filter_data(load_evs_data(), country_param)
}
```

# Overview

This report was generated from the **EVS Data Explorer** Shiny app on `r format(Sys.Date(), "%B %d, %Y")`.

| Parameter | Value |
|---|---|
| Country / Sample | `r country_disp` |
| Outcome variable | `r outcome_label(outcome_param)` |
| Additional controls | `r controls_disp` |
| Age polynomial degree | `r age_poly_param` |
| Regression formula | `r formula_str` |

This report uses two variables of interest found in this EVS 2017 data set. The first variable (v72) is concerned with the level of agreement with the statement: * Men should have more right to a job than women when jobs are scarce.*

The second variable (v80) is concerned with the level of agreement with the statement: * Employers should give priority to nationals over immigrants when jobs are scarce.* 

Looking at these two variables of interest provides valuable information about how these attitudes vary age, sex and education to inform policy making across Europe.
---

# Exploratory Analysis

## Outcome Variable

The chart below shows the distribution of **`r outcome_label(outcome_param)`**
for `r country_disp`.

```{r plot-outcome, fig.height=3.5}
col <- if (outcome_param == "v72") "#2166ac" else "#d6604d"

df %>%
  filter(!is.na(.data[[outcome_param]])) %>%
  count(value = .data[[outcome_param]]) %>%
  mutate(
    pct   = n / sum(n) * 100,
    label = factor(value, levels = 1:5,
                   labels = c("1\nStrongly agree", "2", "3", "4",
                              "5\nStrongly disagree"))
  ) %>%
  ggplot(aes(x = label, y = pct)) +
  geom_col(fill = col, alpha = 0.88, width = 0.7) +
  geom_text(aes(label = sprintf("%.1f%%", pct)), vjust = -0.4, size = 3.5) +
  scale_y_continuous(labels = scales::label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.12))) +
  labs(x = outcome_label(outcome_param), y = "% of respondents") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())
```

## Predictor Variables

### Age

```{r plot-age, fig.height=3}
df %>% filter(!is.na(age)) %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "#4C6A92", colour = "white", alpha = 0.88) +
  labs(x = "Age (years)", y = "Count") +
  theme_minimal(base_size = 12)
```

**Descriptive statistics for age:**

```{r age-summary}
df %>%
  filter(!is.na(age)) %>%
  summarise(N = n(), Mean = round(mean(age), 2),
            SD = round(sd(age), 2), Min = min(age), Max = max(age)) %>%
  kable(caption = "Age descriptive statistics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

### Education Level

```{r plot-edu, fig.height=3}
df %>% filter(!is.na(edu_f)) %>%
  count(edu_f) %>%
  mutate(pct = n / sum(n) * 100) %>%
  ggplot(aes(x = edu_f, y = pct)) +
  geom_col(fill = "#4C6A92", alpha = 0.88, width = 0.6) +
  geom_text(aes(label = sprintf("%.1f%%", pct)), vjust = -0.4, size = 3.5) +
  scale_y_continuous(labels = scales::label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.12))) +
  labs(x = "Education level", y = "% of respondents") +
  theme_minimal(base_size = 12)
```

### Sex

```{r plot-sex, fig.height=3}
df %>% filter(!is.na(sex_f)) %>%
  count(sex_f) %>%
  mutate(pct = n / sum(n) * 100) %>%
  ggplot(aes(x = sex_f, y = pct)) +
  geom_col(fill = "#4C6A92", alpha = 0.88, width = 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", pct)), vjust = -0.4, size = 3.5) +
  scale_y_continuous(labels = scales::label_percent(scale = 1),
                     expand = expansion(mult = c(0, 0.12))) +
  labs(x = "Sex", y = "% of respondents") +
  theme_minimal(base_size = 12)
```

---

# Regression Analysis

## Model

The model below uses OLS regression. The formula is:

**`r formula_str`**

```{r fit-model}
model <- fit_model(df, outcome_param, controls_param, age_poly_param)
```

## Coefficient Table

```{r reg-table}
tidy(model, conf.int = TRUE) %>%
  mutate(
    across(where(is.numeric), ~ round(.x, 4)),
    Sig. = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE            ~ ""
    )
  ) %>%
  rename(Term = term, Estimate = estimate, `Std. Error` = std.error,
         `t value` = statistic, `p value` = p.value,
         `CI 2.5%` = conf.low, `CI 97.5%` = conf.high) %>%
  kable(caption = paste("OLS coefficients —", outcome_label(outcome_param))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE) %>%
  footnote(general = "*** p<0.001  ** p<0.01  * p<0.05  . p<0.1")
```

```{r model-fit}
glance(model) %>%
  select(r.squared, adj.r.squared, sigma, statistic, p.value, nobs) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  rename(`R²` = r.squared, `Adj. R²` = adj.r.squared,
         `Residual SE` = sigma, `F statistic` = statistic,
         `p value` = p.value, N = nobs) %>%
  kable(caption = "Model fit statistics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

## Residuals vs. Fitted Values

```{r plot-resid, fig.height=4.5}
broom::augment(model) %>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.18, colour = "#4C6A92", size = 1.2) +
  geom_hline(yintercept = 0, colour = "#e74c3c", linewidth = 0.8,
             linetype = "dashed") +
  geom_smooth(method = "loess", se = FALSE, colour = "#2C3E50",
              linewidth = 0.8) +
  labs(x = "Fitted values", y = "Residuals",
       title = "Residuals vs. Fitted values") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())
```

---

*Report generated by the EVS Data Explorer · `r format(Sys.time(), "%Y-%m-%d %H:%M")`*
